---
title: "Model simplified reference dataset"
author: Raphael Saldanha
date: last-modified
---

On this notebook, models are trained and tested based on the **simplified version** of the reference dataset.

## Packages

```{r}
#| message: false
library(tidyverse)
library(arrow)
library(qs)
library(tidymodels)
library(themis)
library(finetune)
library(bonsai)
library(tictoc)
library(vip)
library(sessioninfo)
library(doParallel)
source("../../functions.R")

registerDoParallel()
```

## Data load

```{r}
dc_ref_simp <- qread(file = "dc_ref_simp.qs")
```

## Dataset split

```{r}
dc_ref_simp_split <- initial_split(
  data = dc_ref_simp, 
  prop = 3/4, 
  strata = CLASSI_FIN
)

dc_ref_simp_training <- training(dc_ref_simp_split)
dc_ref_simp_testing <- testing(dc_ref_simp_split)
```

## Folding

```{r}
dc_ref_simp_split_folds <- dc_ref_simp_training |>
  vfold_cv(v = 10, strata = CLASSI_FIN)
```

## Recipes

### Variable lists

```{r}
to_dummy <- c('CS_SEXO', 'CS_GESTANT', 'CS_RACA', 'CS_ESCOL_N', 'ID_OCUPA_N', 'HOSPITALIZ')
```

```{r}
symp_cond_vars <- c("FEBRE", "MIALGIA", "CEFALEIA", 
                    "EXANTEMA", "VOMITO", "NAUSEA", 
                    "DOR_COSTAS", "CONJUNTVIT", 
                    "ARTRITE", "ARTRALGIA", "PETEQUIA_N", 
                    "LEUCOPENIA", "LACO", "DOR_RETRO", 
                    "DIABETES", "HEMATOLOG", "HEPATOPAT", 
                    "RENAL", "HIPERTENSA", "ACIDO_PEPT",
                    "AUTO_IMUNE")
```

### Basic recipe

-   Only case symptoms

-   Symptoms with missing data

-   Without date and municipality data

```{r}
rec_ref_simp_1 <- recipe(CLASSI_FIN ~ ., data = dc_ref_simp_training) |>
  step_unknown(all_of(to_dummy)) |>
  step_dummy(all_of(to_dummy)) |>
  step_rm(CRITERIO, case_source, ID_MN_RESI, uf_res, DT_SIN_PRI)
```

### Basic recipe with missing data replacement

-   Only case symptoms

-   Symptoms without missing data

-   Without date and municipality data

```{r}
rec_ref_simp_2 <- recipe(CLASSI_FIN ~ ., data = dc_ref_simp_training) |>
  step_unknown(all_of(to_dummy)) |>
  step_dummy(all_of(to_dummy)) |>
  step_mutate_at(all_of(symp_cond_vars), fn = ~replace_na(., 0)) |>
  step_rm(CRITERIO, case_source, ID_MN_RESI, uf_res, DT_SIN_PRI)
```

### Replace missing recipe

-   Simple recipe and replace symptons NAs with `No` values

```{r}
rec_ref_simp_3 <- recipe(CLASSI_FIN ~ ., data = dc_ref_simp_training) |>
  step_unknown(all_of(to_dummy)) |>
  step_dummy(all_of(to_dummy)) |>
  step_mutate_at(all_of(symp_cond_vars), fn = ~replace_na(., 0)) |>
  step_rm(CRITERIO, case_source, ID_MN_RESI, uf_res, DT_SIN_PRI)
```

### Date recipe

-   Case symptoms and date (month)

-   Without UF

```{r}
rec_ref_simp_4 <- recipe(CLASSI_FIN ~ ., data = dc_ref_simp_training) |>
  step_unknown(all_of(to_dummy)) |>
  step_dummy(all_of(to_dummy)) |>
  step_mutate_at(all_of(symp_cond_vars), fn = ~replace_na(., 0)) |>
  step_rm(CRITERIO, case_source, ID_MN_RESI, uf_res) |>
  step_date(DT_SIN_PRI, features = c("month"), label = FALSE, keep_original_cols = FALSE)
```

### Date and UF recipe

-   Case symptoms, date (month) and UF

```{r}
rec_ref_simp_5 <- recipe(CLASSI_FIN ~ ., data = dc_ref_simp_training) |>
  step_unknown(all_of(to_dummy)) |>
  step_dummy(all_of(to_dummy)) |>
  step_mutate_at(all_of(symp_cond_vars), fn = ~replace_na(., 0)) |>
  step_rm(CRITERIO, case_source, ID_MN_RESI) |>
  step_date(DT_SIN_PRI, features = c("month"), label = FALSE, keep_original_cols = FALSE) |>
  step_dummy(uf_res)
```

## Learners

```{r}
xgb_spec <-
  boost_tree(
    trees = tune(),
    tree_depth = tune(),
    min_n = tune(),
    loss_reduction = tune(),
    learn_rate = tune()
  ) |>
  set_engine("xgboost", event_level = "second") |>
  set_mode("classification")
```

```{r}
# lgbm_spec <-
#   boost_tree(
#     trees = tune(),
#     tree_depth = tune(),
#     min_n = tune(),
#     loss_reduction = tune(),
#     learn_rate = tune()
#   ) |>
#   set_engine("lightgbm") |>
#   set_mode("classification")
```

## Workflow

```{r}
ref_simp_wf <- workflow_set(
  preproc = list(
    rec_ref_simp_1 = rec_ref_simp_1,
    rec_ref_simp_2 = rec_ref_simp_2
    #rec_ref_simp_3 = rec_ref_simp_3,
    #rec_ref_simp_4 = rec_ref_simp_4,
    #rec_ref_simp_5 = rec_ref_simp_5
  ), 
  models = list(xgb = xgb_spec), 
  cross = TRUE
)
```

## Tuning

```{r}
race_ctrl <- control_race(parallel_over = "everything")

ref_simp_tuning_results <- 
  ref_simp_wf |>
  workflow_map(
    "tune_race_anova",
    seed = 345,
    resamples = dc_ref_simp_split_folds,
    grid = 25,
    control = race_ctrl, 
    verbose = TRUE
  )
```

```{r}
ref_simp_tuning_ranking_results <- rank_results(ref_simp_tuning_results, rank_metric = "roc_auc")

ref_simp_tuning_ranking_results
```

```{r}
autoplot(ref_simp_tuning_results, metric = "roc_auc")
```

## Model last fit

```{r}
ref_simp_selection <- ref_simp_tuning_ranking_results |>
  arrange(-mean) |>
  pull(wflow_id) |>
  first()

ref_simp_selection
```

```{r}
ref_simp_best_results <- ref_simp_tuning_results |>
  extract_workflow_set_result(ref_simp_selection) |>
  select_best("roc_auc")

ref_simp_best_results
```

```{r}
ref_simp_last_fit <- ref_simp_tuning_results |>
  extract_workflow(ref_simp_selection) |>
  finalize_workflow(ref_simp_best_results) |>
  last_fit(dc_ref_simp_split)
```

## Model evaluation

```{r}
collect_metrics(ref_simp_last_fit)
```

```{r}
collect_predictions(ref_simp_last_fit) %>%
    conf_mat(CLASSI_FIN, .pred_class)
```

```{r}
ref_simp_last_fit %>%
  extract_fit_engine() %>%
  vip()
```

## Session info

```{r}
session_info()
```
