---
title: "Exported data"
author: Raphael Saldanha
date: last-modified
---

## Packages

```{r}
#| message: false
library(tidyverse)
library(lubridate)
library(arrow)
library(timetk)
library(brpop)
library(piggyback)
```

## Dengue data

Total number of confirmed dengue cases, aggregated per municipality of residence and week of the first symptom's onset.

```{r}
dengue <- read_parquet("../dengue-data/parquet_aggregated/dengue_md.parquet") %>%
  group_by(mun) %>%
  summarise_by_time(.date_var = date, .by = "week", freq = sum(freq, na.rm = TRUE)) %>%
  ungroup() %>%
  rename(cases = freq)
```

## Population

Estimated municipality population per year.

```{r}
pop <- mun_pop_totals() %>%
  filter(year %in% seq(year(min(dengue$date)), year(max(dengue$date)))) %>%
  mutate(mun = as.character(mun))
```

## Human Development Index

```{r}
hdi <- read_parquet("../socioeconomic-data/hdi.parquet") %>%
  select(code_muni, hdi2010 = idhm2010)
```

## Weather data

Weather indicators estimated by using zonal statistics of the territorial area of the municipality.

### Precipitation

Total estimated precipitation, per municipality and week, in millimeter.

```{r}
prec <- open_dataset(sources = "../weather-data/parquet/brdwgd/pr.parquet") %>%
  filter(name == "pr_sum") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue$date) & date <= max(dengue$date)) %>%
  summarise_by_time(.date_var = date, .by = "week", value = sum(value, na.rm = TRUE)) %>%
  rename(prec = value)
```

### Average maximun temperature

Average of maximum temperatures, per municipality and week, in celsius.

```{r}
tmax <- open_dataset(sources = "../weather-data/parquet/brdwgd/tmax.parquet") %>%
  filter(name == "Tmax_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue$date) & date <= max(dengue$date)) %>%
  summarise_by_time(.date_var = date, .by = "week", value = mean(value, na.rm = TRUE)) %>%
  rename(tmax = value)
```

### Average minimum temperature

Average of minimum temperatures, per municipality and week, in celsius.

```{r}
tmin <- open_dataset(sources = "../weather-data/parquet/brdwgd/tmin.parquet") %>%
  filter(name == "Tmin_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue$date) & date <= max(dengue$date)) %>%
  summarise_by_time(.date_var = date, .by = "week", value = mean(value, na.rm = TRUE)) %>%
  rename(tmin = value)
```

## Join data

```{r}
exp_data <- dengue %>%
  mutate(dengue_year = year(date)) %>%
  inner_join(pop, by = c("dengue_year" = "year", "mun")) %>%
  select(-dengue_year) %>%
  inner_join(hdi, by = c("mun" = "code_muni")) %>%
  inner_join(prec, by = "date") %>%
  inner_join(tmax, by = "date") %>%
  inner_join(tmin, by = "date")
```

## Lag weather variables

Lag one and two weeks.

```{r}
exp_data <- exp_data %>%
  tk_augment_lags(.value = c(prec, tmax, tmin), .lags = 1:2)
```

## Overview

```{r}
glimpse(exp_data)
```

Municipalities count: `r length(unique(exp_data$mun))`

## Export data

```{r}
exp_data %>% write_parquet(sink = "exp_data.parquet")
```

```{r}
exp_data %>% write_csv(file = "exp_data.csv")
zip::zip(zipfile = "exp_data.csv.zip", files = "exp_data.csv")
unlink(x = "exp_data.csv")
```

## Data dictionary

-   `mun` character. Municipality code with 6 digits

-   `date` date. Date on format YYYY-MM-DD of the ceiling data of the week

-   `cases` integer. Confirmed dengue cases count

-   `pop` integer. Population estimation of the municipality

-   `hdi2010` double. Human Development Index for 2010

-   `prec` double. Total precipitation, mm

-   `tmax` double. Average maximum temperature, Celsius

-   `tmin` double. Average minimum temperature, Celsius

-   `*_lag1` double. One week lagged variables

-   `*_lag2` double. Two weeks lagged variables

## Include on current GitHub release

```{r}
# Files list to upload
files_list <- c("exp_data.parquet", "exp_data.csv.zip")

# Upload files
for(i in files_list){
  pb_upload(file = i, repo = "rfsaldanha/dengue", overwrite = TRUE)
}
```

Files `exp_data.parquet` and `exp_data.csv.zip` available on current release: <https://github.com/rfsaldanha/dengue/releases>

## Session info

```{r}
sessionInfo()
```
