---
title: "Imputation"
author: Raphael Saldanha
date: last-modified
---

This notebook aims to impute some missing data and enrich the dataset.

## Packages

```{r}
#| message: false
library(tidyverse)
library(arrow)
library(knitr)
library(lubridate)
source("../functions.R")
```

### Execution node

```{r}
node_name()
```

## Load data

```{r}
important_vars <- c("ID_AGRAVO", "DT_NOTIFIC", "ID_UNIDADE",
                    "DT_SIN_PRI", "CS_SEXO", "CS_GESTANT",
                    "CS_RACA", "CS_ESCOL_N", "ID_MN_RESI",
                    "COUFINF", "COMUNINF", "ID_OCUPA_N",
                    "DT_SORO", "RESUL_SORO", "SOROTIPO", 
                    "CLASSI_FIN", "CRITERIO", "EVOLUCAO",
                    "DT_OBITO", "HOSPITALIZ", "DT_INTERNA")

dengue <- open_dataset(sources = data_dir("dengue_data/parquets")) %>%
  select(all_of(important_vars)) %>%
  collect()
```

### Residence municipality: ID_MN_RESI

Task: If ID_MN_RESI is invalid or missing, imputate with valid COMUNINF information.

#### Imputation

```{r}
dengue <- dengue %>%
  mutate(
    ID_MN_RESI_check = if_else(nchar(ID_MN_RESI) == 6, 
                               false = FALSE,
                               true = TRUE),
    COMUNINF_check = if_else(nchar(COMUNINF) == 6, 
                               false = FALSE,
                               true = TRUE)
  ) %>%
  mutate(ID_MN_RESI = case_when(
    ID_MN_RESI_check == FALSE & COMUNINF_check == TRUE ~ COMUNINF,
    is.na(ID_MN_RESI) ~ COMUNINF,
    TRUE ~ ID_MN_RESI
  )) %>%
  select(-ID_MN_RESI_check, -COMUNINF_check)
```

#### Check improvement

```{r}
dengue %>%
  mutate(
    ID_MN_RESI_check = if_else(nchar(ID_MN_RESI) == 6, 
                               false = FALSE,
                               true = TRUE)
  ) %>%
  group_by(ID_MN_RESI_check) %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

No improvement for invalid municipalities, improvement of 96 records with missing data.

### Date of the first symptoms onset: DT_SIN_PRI

Task: If DT_SIN_PRI is invalid or missing, imputate with valid DT_NOTIFIC information.

#### Imputation

```{r}
valid_interval <- interval(ymd("2011-01-01"), ymd("2023-12-31"))

dengue <- dengue %>%
  mutate(
    DT_SIN_PRI_check = ymd(DT_SIN_PRI) %within% valid_interval,
    DT_NOTIFIC_check = ymd(DT_NOTIFIC) %within% valid_interval,
  ) %>%
  mutate(DT_SIN_PRI = case_when(
    DT_SIN_PRI_check == FALSE & DT_NOTIFIC_check == TRUE ~ DT_NOTIFIC,
    is.na(DT_SIN_PRI) ~ DT_NOTIFIC,
    TRUE ~ DT_SIN_PRI
  )) %>%
  select(-DT_SIN_PRI_check, -DT_NOTIFIC_check)
```

#### Check improvement

```{r}
valid_interval <- interval(ymd("2011-01-01"), ymd("2023-12-31"))

dengue %>%
  mutate(
    DT_SIN_PRI_check = ymd(DT_SIN_PRI) %within% valid_interval
  ) %>%
  group_by(DT_SIN_PRI_check) %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

There was a improvement on 25,823 records with invalid dates and all missing records were imputate.

## Export improved database

```{r}
dengue %>%
  write_parquet(sink = data_dir("dengue_data/parquet_improved/dengue_improved.parquet"))
```

## Session info

```{r}
sessionInfo()
```
