---
title: "Missing data analysis"
author: Raphael Saldanha
date: last-modified
---

This report presents a missing data analysis from the raw parquet files.

Considering the high number of variables, this report will use a subset of the most relevant ones for this research.

## Packages

```{r}
#| message: false
library(tidyverse)
library(arrow)
library(naniar)
library(knitr)
library(lubridate)
source("../functions.R")
```

### Execution node

```{r}
node_name()
```

## Load data

```{r}
important_vars <- c("ID_AGRAVO", "DT_NOTIFIC", "ID_UNIDADE",
                    "DT_SIN_PRI", "CS_SEXO", "CS_GESTANT",
                    "CS_RACA", "CS_ESCOL_N", "ID_MN_RESI",
                    "COUFINF", "COMUNINF", "ID_OCUPA_N",
                    "DT_SORO", "RESUL_SORO", "SOROTIPO", 
                    "CLASSI_FIN", "CRITERIO", "EVOLUCAO",
                    "DT_OBITO", "HOSPITALIZ", "DT_INTERNA")

dengue <- open_dataset(sources = file.path(c(data_dir(), "parquets/"))) %>%
  select(all_of(important_vars)) %>%
  collect()
```

## Overall

Considering all records.

```{r}
dengue %>% 
  miss_var_summary() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

### Variables quality

#### Residence municipality: ID_MN_RESI

Check var length. 6 characters are expected.

```{r}
dengue %>%
  mutate(
    ID_MN_RESI_check = if_else(nchar(ID_MN_RESI) == 6, 
                               false = FALSE,
                               true = TRUE)
  ) %>%
  group_by(ID_MN_RESI_check) %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )


```

15.559.994 records on the database meets the criteria. 4 records have invalid municipalty codes and this information is missing for 2.126 records. Inputation from COMUNINF is a possibility.

#### Final classification of the notification: CLASSI_FIN

It is expected that the notifications final classification are labeled.

```{r}
dengue %>%
  group_by(CLASSI_FIN) %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

3 records have an invalid label and 69.866 are missing. This missing data may be from two causes: (1) the notification is being evaluated or (2) a real missing.

#### Date of the first symptons onset: DT_SIN_PRI

This is the date most near to the infection date and more relevant to epidemiological analysis.

```{r}
valid_interval <- interval(ymd("2001-01-01"), ymd("2021-12-31"))

dengue %>%
  mutate(
    DT_SIN_PRI_check = ymd(DT_SIN_PRI) %within% valid_interval
  ) %>%
  group_by(DT_SIN_PRI_check) %>%
  summarise(freq = n()) %>%
  ungroup() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

This variable is missing for 173 records. For 26.018 records, we have invalid dates (outside the period 2001-2021). It is possible to imputate with DT_NOTIFIC (date of notification), if valid.

## Positive cases

```{r}
dengue_classifications <- c("Febre hemorrágica do dengue", "Síndrome do choque do dengue", "Dengue com sinais de alarme",
  "Dengue clássico", "Dengue com complicações", "Dengue",
  "Dengue grave")

dengue %>% 
  filter(CLASSI_FIN %in% dengue_classifications) %>%
  miss_var_summary() %>%
  kable(
    format.args = list(big.mark = ".", decimal.mark = ",")
  )
```

## Session info

```{r}
sessionInfo()
```
