---
title: "Global and subset models"
author: "Raphael Saldanha"
date: last-modified
---

Reference: https://blog.bguarisma.com/series/time-series-forecasting

## Packages

```{r}
#| message: false
library(tidyverse)
library(arrow)
library(tidymodels)
library(finetune)
library(modeltime)
library(timetk)
library(tictoc)
```

## Load data

And remove rows with NAs

```{r}
tdengue <- read_parquet(file = "tdengue.parquet") %>%
  drop_na()
```

## Cluster

Assign random groups.

::: callout-warning
To use a cluster algorithm later.
:::

```{r}
tdengue <- tdengue %>%
  mutate(group = sample(x = 4, size = nrow(.), replace = TRUE))
```

## Train and test split

```{r}
tdengue_split <- tdengue %>%
  time_series_split(
    date_var = date, 
    assess = 24,
    cumulative = TRUE
  )
```

## K-folds

```{r}
tdengue_split_folds <- training(tdengue_split) %>%
  vfold_cv(v = 5)
```

## Recipes

```{r}
recipes_list <- list()
```

### Global

Using data from all municipalities.

```{r}
recipe_global <- recipe(cases ~ ., data = training(tdengue_split)) %>%
  step_rm(mun, group)

recipes_list <- append(recipes_list, list(global = recipe_global))

rm(recipe_global)
```

### Groups

One recipe for each group, with data only from the group.

```{r}
for(g in unique(tdengue$group)){
  tmp <- recipe(cases ~ ., data = training(tdengue_split)) %>%
  step_filter(group == g) %>%
  step_rm(mun, group)
  
  tmp <- list(tmp)
  tmp <- setNames(tmp, paste0("g", g))
  
  recipes_list <- append(recipes_list, tmp)
  
  rm(tmp)
}
```

## Models specification

### Random forest

```{r}
rf_spec <- rand_forest(
  trees = tune(),
  min_n = tune()
) %>%
  set_engine("ranger") %>%
  set_mode("regression")
```

### Decision tree

```{r}
dt_spec <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("regression")
```

## Workflow set

```{r}
all_workflows <- workflow_set(
  preproc = recipes_list, 
  models = list(rf = rf_spec, dt = dt_spec), 
  cross = TRUE
)
```

## Tune

```{r}
doParallel::registerDoParallel()

tic()
race_results <- 
  all_workflows %>%
  workflow_map(
    "tune_race_anova",
    seed = 345,
    resamples = tdengue_split_folds,
    grid = 3,
    control = control_race(parallel_over = "everything"),
    verbose = TRUE
  )
toc()
```

## Fit train data

For combination of each municipality, model and recipe.

```{r}
trained_models <- list()
for(w in unique(race_results$wflow_id)){
  best_tune <- race_results %>%
    extract_workflow_set_result(w) %>%
    select_best("rmse")
  
  final_fit <- race_results %>%
    extract_workflow(w) %>%
    finalize_workflow(best_tune) %>%
    fit(training(tdengue_split))
  
  final_fit <- list(final_fit)
  final_fit <- setNames(final_fit, paste0(w))
  
  trained_models <- append(trained_models, final_fit)
}
```

```{r}
models_accuracy <- tibble()
for(t in 1:length(trained_models)){
  model_tbl <- modeltime_table(trained_models[[t]]) 

  calib_tbl <- model_tbl %>%
      modeltime_calibrate(
        new_data = testing(tdengue_split), 
        id       = "mun"
      )
  
  res <- calib_tbl %>% 
      modeltime_accuracy(
        acc_by_id = TRUE, 
        metric_set = metric_set(rmse)
      )
  
  res$.model_id <- word(names(trained_models[t]), 1, sep = "_")
  
  models_accuracy <- bind_rows(models_accuracy, res)
}
```

```{r}
ggplot(data = models_accuracy, aes(x = .model_id, y = rmse, fill = .model_desc)) +
  geom_boxplot()
```

## Session info

```{r}
sessionInfo()
```
