---
title: "Dengue case classification"
subtitle: "by symptoms and clinical condition"
author: Raphael Saldanha
date: last-modified
---

The objective of this notebook is to train a model to reclassify inconclusive cases of dengue, zika and chikungunya based on the symptoms, clinical conditions and other patient related variables presented by confirmed and discarded cases

## Packages

```{r}
#| message: false
library(tidyverse)
library(arrow)
library(knitr)
library(lubridate)
library(tidymodels)
library(finetune)
library(bonsai)
library(tictoc)
library(vip)
source("../functions.R")
```

## Data

-   Cases classified as inconclusive are discarded for model training.

### Dengue

-   Data prior to 2016 does not have patient symptoms and clinical conditions.

```{r}
# Data sources
files_list <- c(
  data_dir("dengue_data/parquets/dengue_2016.parquet"),
  data_dir("dengue_data/parquets/dengue_2017.parquet"),
  data_dir("dengue_data/parquets/dengue_2018.parquet"),
  data_dir("dengue_data/parquets/dengue_2019.parquet"),
  data_dir("dengue_data/parquets/dengue_2020.parquet"),
  data_dir("dengue_data/parquets/dengue_2021.parquet")
)

# Independent variables
x_vars <- c("FEBRE", "MIALGIA", "CEFALEIA", 
            "EXANTEMA", "VOMITO", "NAUSEA", 
            "DOR_COSTAS", "CONJUNTVIT", 
            "ARTRITE", "ARTRALGIA", "PETEQUIA_N", 
            "LEUCOPENIA", "LACO", "DOR_RETRO", 
            "DIABETES", "HEMATOLOG", "HEPATOPAT", 
            "HEPATOPAT", "RENAL", "HIPERTENSA",
            "ACIDO_PEPT", "AUTO_IMUNE")

# Prepare data
dengue <- arrow::open_dataset(sources = files_list) %>%
  # Select variables
  select(all_of(c("CLASSI_FIN", "COMUNINF", "IDADEanos", "DT_SIN_PRI", x_vars))) %>%
  # Filter out "Inconclusivo" cases
  filter(CLASSI_FIN != "Inconclusivo") %>%
  # Collect data from parquet files
  collect() %>%
  # Prepare variables
  mutate(CLASSI_FIN = case_when(
    CLASSI_FIN != "Descartado" ~ "Dengue",
    .default = "Discarded"
  )) %>%
  mutate(CLASSI_FIN = as.factor(CLASSI_FIN)) %>%
  mutate(DT_SIN_PRI = as_date(DT_SIN_PRI)) %>%
  mutate(COMUNINF = as.factor(COMUNINF)) %>%
  mutate_at(.vars = x_vars, .funs = ~ . == "Sim") 
```


### Chikungunya

-   Data prior to 2017 does not have patient symptoms and clinical conditions.

```{r}
# Data sources
files_list <- c(
  data_dir("chik_data/parquets/chik_2017.parquet"),
  data_dir("chik_data/parquets/chik_2018.parquet"),
  data_dir("chik_data/parquets/chik_2019.parquet"),
  data_dir("chik_data/parquets/chik_2020.parquet"),
  data_dir("chik_data/parquets/chik_2021.parquet")
)

# Independent variables
x_vars <- c("FEBRE", "MIALGIA", "CEFALEIA", 
            "EXANTEMA", "VOMITO", "NAUSEA", 
            "DOR_COSTAS", "CONJUNTVIT", 
            "ARTRITE", "ARTRALGIA", "PETEQUIA_N", 
            "LEUCOPENIA", "LACO", "DOR_RETRO", 
            "DIABETES", "HEMATOLOG", "HEPATOPAT", 
            "HEPATOPAT", "RENAL", "HIPERTENSA",
            "ACIDO_PEPT", "AUTO_IMUNE")

# Prepare data
chik <- arrow::open_dataset(sources = files_list) %>%
  # Select variables
  select(all_of(c("CLASSI_FIN", "COMUNINF", "IDADEanos", "DT_SIN_PRI", x_vars))) %>%
  # Filter out "Inconclusivo" cases
  filter(CLASSI_FIN != "Inconclusivo") %>%
  # Collect data from parquet files
  collect() %>%
  # Prepare variables
  mutate(CLASSI_FIN = case_when(
    CLASSI_FIN != "Descartado" ~ "Chikungunya",
    .default = "Discarded"
  )) %>%
  mutate(CLASSI_FIN = as.factor(CLASSI_FIN)) %>%
  mutate(DT_SIN_PRI = as_date(DT_SIN_PRI)) %>%
  mutate(COMUNINF = as.factor(COMUNINF)) %>%
  mutate_at(.vars = x_vars, .funs = ~ . == "Sim") 
```

### Work dataset

```{r}
dcdata <- bind_rows(dengue, chik)
```

```{r}
rm(dengue, chik)
gc(full = TRUE, verbose = FALSE)
```

## Case classification distribution

```{r}
dim(dcdata)
```


```{r}
dcdata %>%
  group_by(CLASSI_FIN) %>%
  summarise(freq = n()) %>%
  ungroup()
```



## Modeling

### Train and test dataset split

-   Proportion between training and test: 3/4

```{r}
set.seed(123)

dcdata_split <- dcdata %>%
  initial_split(prop = 3/4, strata = CLASSI_FIN)

dcdata_train <- training(dcdata_split)
dcdata_test <- testing(dcdata_split)

dcdata_folds <- vfold_cv(dcdata_train, v = 10, strata = CLASSI_FIN)
```

### Recipes

```{r}
rec_1 <- 
  recipe(CLASSI_FIN ~ . , data = dcdata_train) %>%
  step_rm(COMUNINF) %>%
  step_rm(DT_SIN_PRI) %>%
  #step_date(DT_SIN_PRI, features = c("month", "week", "semester", "quarter"), keep_original_cols = FALSE) %>%
  step_integer(all_predictors())
```

### Specifications

#### XGB

```{r}
xgb_spec <-
  boost_tree(
    trees = tune(),
    min_n = tune(),
    mtry = tune()
  ) %>%
  set_engine("xgboost") %>%
  set_mode("classification")
```

#### Decision tree

```{r}
cart_spec <-
  decision_tree(
    cost_complexity = tune(), 
    min_n = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")
```

### Workflows

```{r}
all_workflows <- 
  workflow_set(
    preproc = list(recipe_1 = dengue_rec_1),
    models = list(xgb = xgb_spec, cart = cart_spec)
  )
```

### Tuning

```{r}
doParallel::registerDoParallel()

race_ctrl <- control_race(parallel_over = "everything")

tic()
race_results <- 
  all_workflows %>%
  workflow_map(
    "tune_race_anova",
    seed = 345,
    resamples = dengue_folds,
    grid = 10,
    control = race_ctrl
  )
toc()
```

#### Race metrics

```{r}
train_rank_results <- rank_results(race_results, rank_metric = "roc_auc")

train_rank_results
```

```{r}
autoplot(race_results, metric = "roc_auc")
```

### Last fit

```{r}
selection_train <- train_rank_results %>%
  arrange(-mean) %>%
  pull(wflow_id) %>%
  first()

selection_train
```

```{r}
best_results <- race_results %>%
  extract_workflow_set_result(selection_train) %>%
  select_best("accuracy")

best_results
```

```{r}
last_fit <- race_results %>%
  extract_workflow(selection_train) %>%
  finalize_workflow(best_results) %>%
  last_fit(dengue_split)
```

### Evaluate on test

```{r}
collect_metrics(last_fit)
```

```{r}
collect_predictions(last_fit) %>%
    conf_mat(CLASSI_FIN, .pred_class)
```

```{r}
last_fit %>%
  extract_fit_engine() %>%
  vip()
```

## Session info

```{r}
sessionInfo()
```
