---
title: "Classification task"
author: Raphael Saldanha
date: last-modified
---

This notebook aims to investigate the relationship between weather variables and dengue cases in different time lags in a classification task.

Initially, the Rio de Janeiro, RJ municipality is examined.

## Packages

```{r}
#| message: false
library(tidyverse)
library(tidymodels)
library(arrow)
library(timetk)
library(rpart.plot)
library(vip)
```

## Dengue data

### Subset and aggregate

Rio de Janeiro, RJ, aggregated by month.

```{r}
dengue_rj <- open_dataset("../dengue-data/parquet_aggregated/dengue_md.parquet") %>%
  filter(mun == 330455) %>%
  collect() %>%
  summarise_by_time(.date_var = date, .by = "month", freq = sum(freq, na.rm = TRUE))
```

```{r}
plot_time_series(.data = dengue_rj, .date_var = date, .value = freq, .smooth = FALSE, .title = "Dengue, absolute number of cases")
```

### Classify

Based on the nominal dengue incidence, we will classify the incidence as outbreak or normal levels. For this, an time series anomaly detection algorithm was applied.

```{r}
plot_anomaly_diagnostics(.data = dengue_rj, .date_var = date, .value = freq, .alpha = 0.10, .max_anomalies = 1, .legend_show = FALSE)
```

```{r}
dengue_rj_anom <- tk_anomaly_diagnostics(.data = dengue_rj, .date_var = date, .value = freq, .alpha = 0.10, .max_anomalies = 1) %>%
  select(date, anomaly) %>%
  mutate(anomaly = as.factor(anomaly))
```

```{r}
dengue_rj <- inner_join(dengue_rj, dengue_rj_anom) 
```

```{r}
prop.table(table(dengue_rj$anomaly))
```

## Weather data

Load weather data and subset to the selected municipality and time period.

```{r}
tmax <- open_dataset(sources = "../weather-data/parquet/brdwgd/tmax.parquet") %>%
  filter(code_muni == 3304557) %>%
  filter(name == "Tmax_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue_rj$date) & date <= max(dengue_rj$date)) %>%
  summarise_by_time(.date_var = date, .by = "month", value = mean(value, na.rm = TRUE)) %>%
  rename(tmax = value)

prec <- open_dataset(sources = "../weather-data/parquet/brdwgd/pr.parquet") %>%
  filter(code_muni == 3304557) %>%
  filter(name == "pr_sum") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue_rj$date) & date <= max(dengue_rj$date)) %>%
  summarise_by_time(.date_var = date, .by = "month", value = sum(value, na.rm = TRUE)) %>%
  rename(prec = value)

prec_avg <- open_dataset(sources = "../weather-data/parquet/brdwgd/pr.parquet") %>%
  filter(code_muni == 3304557) %>%
  filter(name == "pr_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue_rj$date) & date <= max(dengue_rj$date)) %>%
  summarise_by_time(.date_var = date, .by = "month", value = mean(value, na.rm = TRUE)) %>%
  rename(prec_avg = value)

rh <- open_dataset(sources = "../weather-data/parquet/brdwgd/rh.parquet") %>%
  filter(code_muni == 3304557) %>%
  filter(name == "RH_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue_rj$date) & date <= max(dengue_rj$date)) %>%
  summarise_by_time(.date_var = date, .by = "month", value = mean(value, na.rm = TRUE)) %>%
  rename(rh = value)

wind <- open_dataset(sources = "../weather-data/parquet/brdwgd/u2.parquet") %>%
  filter(code_muni == 3304557) %>%
  filter(name == "u2_mean") %>%
  select(date, value) %>%
  collect() %>%
  filter(date >= min(dengue_rj$date) & date <= max(dengue_rj$date)) %>%
  summarise_by_time(.date_var = date, .by = "month", value = mean(value, na.rm = TRUE)) %>%
  rename(wind = value)
```

```{r}
plot_time_series(.data = tmax, .date_var = date, .value = tmax, .smooth = FALSE, .title = "Max temp, average")
```

```{r}
plot_time_series(.data = prec, .date_var = date, .value = prec, .smooth = FALSE, .title = "Precipitation, sum")
```

```{r}
plot_time_series(.data = prec_avg, .date_var = date, .value = prec_avg, .smooth = FALSE, .title = "Precipitation, average")
```

```{r}
plot_time_series(.data = rh, .date_var = date, .value = rh, .smooth = FALSE, .title = "Relative humidity, average")
```

```{r}
plot_time_series(.data = wind, .date_var = date, .value = wind, .smooth = FALSE, .title = "Wind, average")
```

## Join data

Joins the dengue incidence classification and weather variables.

```{r}
res <- inner_join(x = dengue_rj, y = tmax, by = "date") %>%
  inner_join(prec, by = "date") %>%
  inner_join(prec_avg, by = "date") %>%
  inner_join(rh, by = "date") %>%
  inner_join(wind, by = "date") %>%
  select(date, anomaly, tmax, prec, rh, wind)
```

## Decision tree

### Prepare data

-   Remove date

-   Lag variables: 12 months

```{r}
res_prep <- res %>%
  select(-date) %>%
  tk_augment_lags(.value = c(tmax, prec, rh, wind), .lags = 1:6)
```

### Parameters

```{r}
tree_spec <- decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")
```

## Fit model

```{r}
fit1 <- tree_spec %>%
  fit(anomaly ~ ., data = res_prep, model = TRUE)
```

```{r}
fit1 %>% extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)
```

```{r}
augment(fit1, new_data = res_prep) %>%
    conf_mat(truth = anomaly, estimate = .pred_class)
```

```{r}
augment(fit1, new_data = res_prep) %>%
    accuracy(truth = anomaly, estimate = .pred_class)
```

```{r}
fit1 %>% 
  extract_fit_engine() %>% 
  vip()
```

## Session info

```{r}
sessionInfo()
```
