---
title: "Cumulative time series"
author: Raphael Saldanha
date: last-modified
---

This notebook aims to cluster the Brazilian municipalities cumulative yearly dengue cases time-series yearly by its similarities.

## Packages

```{r}
#| message: false
library(tidyverse)
library(lubridate)
library(arrow)
library(timetk)
library(dtwclust)
library(kableExtra)
library(tictoc)
source("../functions.R")
```

## Load data

Load the bundled data.

```{r}
dengue <- open_dataset(sources = data_dir("bundled_data/tdengue.parquet")) %>%
    select(mun, date, cases = cases_raw) %>%
    collect()

dim(dengue)
```

### Prepare data

The chunk bellow executes various steps to prepare the data for clustering.

```{r}
tdengue <- dengue %>%
  # Remove dates
  filter(date >= as.Date("2011-01-01") & date < as.Date("2020-01-01")) %>%
  # Create year variable
  mutate(year = year(date)) %>%
  # Create week number
  mutate(week = epiweek(date)) %>%
  # Summarise per year and isoweek
  group_by(mun, year, week) %>%
  summarise(cases = sum(cases, na.rm = TRUE)) %>%
  # Accumulate mun time series by year
  mutate(freq = cumsum(cases)) %>%
  # Scale cases
  mutate(cases = scale(cases)) %>%
  ungroup() %>%
  # Isolate municipality and year
  mutate(mun = paste0(mun, "_", year)) %>%
  select(-year, week) %>%
  # Arrange data
  arrange(mun, week) %>%
  # Prepare time series of municipalities by year
  mutate(mun = paste0("m_", mun)) %>%
  pivot_wider(names_from = mun, values_from = cases) %>%
  select(-week) %>%
  # Use zero value for years withot week 53
  mutate(across(everything(), ~replace_na(.x, 0))) %>%
  # Transpose as matrix
  t() %>%
  # Convert object
  tslist()
```

```{r}
length(tdengue)
```

## Clustering

Sequence of `k` groups to be used.

```{r}
k_seq <- 3:10
```

### Soft-DTW method

```{r}
tic()
clust <- tsclust(
  series = tdengue, 
  type = "partitional", 
  k = k_seq,
  distance = "sdtw",
  seed = 13
)
toc()
```

### Cluster Validity Indices (CVI)

```{r}
names(clust) <- paste0("k_", k_seq)
res_cvi <- sapply(clust, cvi, type = "internal") %>% 
  t() %>%
  as_tibble(rownames = "k") %>%
  arrange(-Sil)

res_cvi %>%
  kbl() %>%
  kable_styling()
```

### Cluster with higher Silhouette statistic

```{r}
sel_clust <- clust[[res_cvi[[1,1]]]]
```

```{r}
plot(sel_clust)
```

```{r}
ggsave(filename = "cluster_mun_year_cum.pdf")
```

```{r}
plot(sel_clust, type = "centroids", lty = 1)
```

```{r}
ggsave(filename = "cluster_mun_year_cum_centr.pdf")
```

### Cluster sizes

```{r}
table(sel_clust@cluster)
```

## Session info

```{r}
sessionInfo()
```
