---
title: "Mun/year time series, cumulative"
author: Raphael Saldanha
date: now
---

This notebook aims to cluster the Brazilian municipalities cumulative dengue cases time-series yearly by its similarities.

## Packages

```{r}
library(tidyverse)
library(lubridate)
library(arrow)
library(timetk)
library(dtwclust)
library(kableExtra)
library(tictoc)
```

## Load data

Load the aggregated data.

```{r}
dengue <- read_parquet("../dengue-data/parquet_aggregated/dengue_md.parquet")

dim(dengue)
```

### Prepare data

The chunk bellow executes various steps to prepare the data for clustering.

```{r}
tdengue <- dengue %>%
  # Filter out dates after 2020-01-01, as this year is not complete
  filter(date <= as.Date("2020-01-01")) %>%
  # Pad days with zero cases
  group_by(mun) %>%
  pad_by_time(date, .by = "day", .pad_value = 0, .start_date = min(dengue$date), .end_date = max(dengue$date)) %>%
  ungroup() %>%
  # Aggregate by week
  mutate(
    week = str_pad(epiweek(date), pad = 0, width = 2),
    year = year(date)
  ) %>%
  group_by(mun, year, week) %>%
  summarise(freq = sum(freq, na.rm = TRUE)) %>%
  ungroup() %>%
  # Keep only municipalities/year series with more than 100 cases total
  group_by(mun, year) %>%
  mutate(total = sum(freq, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(total >= 100) %>%
  select(-total) %>%
  # Modify freq
  group_by(mun, year) %>%
  arrange(week) %>%
  # Accumulate mun time series by year
  mutate(freq = cumsum(freq)) %>%
  # Scale cases by mun and year
  mutate(freq = scale(freq)) %>%
  ungroup() %>%
  arrange(mun, year, week) %>%
  # Isolate municipality and year
  mutate(mun = paste0(mun, "_", year)) %>%
  select(-year, week) %>%
  # Prepare time series of municipalities by year
  mutate(mun = paste0("m_", mun)) %>%
  arrange(mun) %>%
  pivot_wider(names_from = mun, values_from = freq) %>%
  select(-week) %>%
  t() %>%
  # Convert object
  tslist()
# Remove missing values created by leap years (week 53)
tdengue <- lapply(tdengue, na.omit)
```

```{r}
length(tdengue)
```

## Clustering

Sequence of `k` groups to be used.

```{r}
k_seq <- 3:10
```

### SBD method

```{r}
tic()
clust <- tsclust(
  series = tdengue, 
  type = "partitional", 
  k = k_seq,
  distance = "sbd",
  seed = 13
)
toc()
```

### Cluster Validity Indices (CVI)

```{r}
names(clust) <- paste0("k_", k_seq)
res_cvi <- sapply(clust, cvi, type = "internal") %>% 
  t() %>%
  as_tibble(rownames = "k") %>%
  arrange(-Sil)

res_cvi %>%
  kbl() %>%
  kable_styling()
```

### Cluster with higher Silhouette statistic

```{r}
sel_clust <- clust[[res_cvi[[1,1]]]]
```

```{r}
plot(sel_clust)
```

```{r}
ggsave(filename = "cluster_mun_year_cum.pdf")
```

```{r}
plot(sel_clust, type = "centroids", lty = 1)
```

### Cluster sizes

```{r}
table(sel_clust@cluster)
```

## Session info

```{r}
sessionInfo()
```
